---
title: "Rita Li FP1"
author: "Rita Li"
date: '2022-09-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Area
Public Transportation

## Members
Jenny, Pippa, Rita


## Research question
- Where are there gaps in a public transportation system?
- What are the connections between public transport systems and    environmental protection and emissions? 
- How much money the government spends (or future spends) on transportation based on energy cost ?

## Dataset
```{r}
library(tidyverse)     # for graphing and data cleaning
library(lubridate)     # for date manipulation
library(ggthemes)      # for even more plotting themes
library(dplyr)
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(readr)
library(readxl)
library(plotly)
Sys.setlocale("LC_TIME", "English")
theme_set(theme_minimal())
```


# Gas Price dataset
```{r}
gas <- read_csv("FP1/Rita-Li-FP1_datasets/Weekly_Minnesota_Regular_Conventional_Retail_Gasoline_Prices.csv", skip = 4)
```
## Data source
[U.S. Energy Information Administration ](https://www.eia.gov/dnav/pet/hist/LeafHandler.ashx?n=PET&s=EMM_EPMRU_PTE_SMN_DPG&f=W)
## Data description
This is the dataset of weekly Minnesota regular conventional retail gasoline price.

## Data limitations (eg: are the data recent? do they contain all variables you might want?)
This dataset only have two variables: time and the price.
## Visualize one interesting variable.


# Metro Transit Ridership Dataset
```{r}
ridership2013 <- read_csv("FP1/Rita-Li-FP1_datasets/TransitStopsBoardingsAndAlightings2013.csv") %>% 
  filter(Provider == "Metro Transit")
```
## Data source
[Metropolitan Council](https://gisdata.mn.gov/dataset/us-mn-state-metc-trans-stop-boardings-alightings)

## Data description
Contains average daily transit boardings and alightings for all Metro Transit, Met Council, and Maple Grove Transit fixed-route transit service.

## Data limitations 
This only contains data from 2013 to 2021


## Visualize one interesting variable.

## Join datasets




